Finding Concurrency Bugs in Production

### 2022-05-15 src/paper-skill.tex
add SKILL entry
Signed-off-by: Marvin Strangfeld <marvin@strangfeld.io>

% ------------------------------------ %
% Copyright (C) 2020 Marvin Strangfeld %
% ------------------------------------ %
\documentclass[conference]{IEEEtran}

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{listings}
\usepackage{xcolor}
\usepackage[hidelinks]{hyperref}
\usepackage{cleveref}
\usepackage{float}

@@

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Finding Concurrency Bugs in Production}

\author{\IEEEauthorblockN{Marvin Alexander Strangfeld}
\IEEEauthorblockA{
    \textit{RWTH Aachen University}\\
    marvin.strangfeld@rwth-aachen.de}
}

\maketitle

% ------------------------------------ %
% ------------- ABSTRACT ------------- %
% ------------------------------------ %
\begin{abstract}
Bugs that are caused by the parallel execution of a program, so called
concurrency bugs are often hard to find and therefore hard to fix.
In this paper we evaluate different techniques and tools that are currently available to effectively find, fix and avoid concurrency bugs.
These tools are evaluated by their ease of use, their computational and storage overhead, their concurrency bug coverage and their potential amount of false-positive reports.
The goal is to find tools that are production-ready and applicable to real-world software projects.
The techniques evaluated are:
\begin{itemize}
    \item Dynamic code analysis in form of record and replay and dynamic data race detection.
    \item Concurrency-aware testing by evaluating thread schedules with delta-debugging to automatically pinpoint concurrency bugs.
    \item Static code analysis with the methods of sequentialization and model-checking.
    \item Programming language design patterns to avoid concurrency bugs when writing software.
\end{itemize}
\end{abstract}

\begin{IEEEkeywords}
    [Software Engineering]: Software testing and Debugging
    [Computing methodologies]: Concurrent programming languages
\end{IEEEkeywords}

% ------------------------------------ %
% ----------- INTRODUCTION ----------- %
% ------------------------------------ %
\section{Introduction}

@@

\subsection{The Go Programming Language}
The increasing demand for parallel programs has also lead to the development of new programming languages which specialize in this task.
The goal of these languages is to ease the usage of parallel programming patterns, so developers can use them more often without writing extensively more code.
Additionally, they try to avoid the creation of concurrency bugs by their design of language which will be evaluated in \Cref{sct:outlook}.
One example for such a language is Go which is a statically typed programming language developed by Google, that compiles down to native binaries~\cite{goDocs}.
Go recently gained a lot of popularity for the purpose of writing highly concurrent programs.
Popular open-source projects like \emph{Kubernetes}\footnote{\url{https://kubernetes.io/}}, \emph{Docker}\footnote{\url{https://www.docker.com/}} and \emph{Prometheus}\footnote{\url{https://prometheus.io/}} are all written in Go.
The official documentation of Go states: ``[... Go's] concurrency mechanisms make it easy to write programs that get the most out of multicore and networked machines [...]''~\cite{goDocs}.
Go abstracts the creation of threads by using lightweight \emph{goroutines}, which can be created by using the \lstinline{go} keyword followed by a named or anonymous function.
The Go runtime maps these goroutines to operating system threads during execution.
Go encourages programmers to use concurrency by \emph{message passing}, which is assumed to be safer and more convenient to use than \emph{shared memory}.
For this purpose Go is using a concept called \emph{channels} where typed messages can be passed through like a UNIX pipeline.
Beside the message passing however, it is also possible to make use of \emph{shared memory}, which gives the developer a lot of freedom to orchestrate the parallel execution of threads.
Even though Go is designed for programming concurrent programs, this does not mean that Go programs are free of concurrency bugs as we will see in \Cref{sct:outlook}.
Projects written in Go also tend to have more concurrency than projects written in traditional programming languages such as C~\cite{tu2019go}.
So it is even more important for Go developers to catch concurrency bugs effectively and as early as possible.
Together with the increasing popularity of the language, the demand for production-ready tools to catch concurrency bugs is very high, which is why this paper evaluates existing techniques in the context of the Go programming language.

@@

The further structure of this paper continues as follows:
\Cref{sct:taxonomy} gives a brief introduction on the different types of concurrency bugs and their main causes.
\Cref{sct:dynamic} covers some techniques to dynamically detect concurrency bugs and reliably reproduce them.
\Cref{sct:testing} shows different methods of concurrency-aware testing to detect concurrency bugs automatically in a testing environment, where in contrast to a production environment concurrency bugs are explicitly provoked.
\Cref{sct:static} covers how to detect concurrency bugs with static code analysis.
And finally, \Cref{sct:outlook} will provide a quick outlook on how the examined tools can be used together as well as other promising technologies that might help avoiding concurrency bugs in the future.

@@

\begin{figure}
    \includegraphics[width=\linewidth]{figures/ConcurrencyBugClasses.pdf}
    \caption{A Taxonomy of Concurrency Bugs -- based on\cite{tchamgoue2012testing}}
    \label{fig:classes}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=\linewidth]{figures/Concurrency-Testing.pdf}
    \caption{Automatic concurrency-aware testing for Go programs -- based on\cite{acm2002}}
    \label{fig:testing}
\end{figure}

@@

% ------------------------------------ %
% -------------- OUTLOOK ------------- %
% ------------------------------------ %
\section{Outlook}
\label{sct:outlook}

All the techniques shown can find concurrency bugs with different coverage based on the overhead one is willing to pay.
As seen, there is not a single tool that fulfills all the requirements for an optimal production usage.
Either the overhead is very big or the coverage is quite small.
To get the most out of these tools, a combined approach seems to make sense to get the most coverage with the least overhead.

For the goal of having no concurrency bugs at all in a production deployment with nearly no overhead, it would be desirable to have techniques that make it impossible to even write incorrect concurrent programs.
New languages and language features have been developed for this task.
Go as presented before is one of those languages that tries to make the use of concurrent patterns safer.
One of their main approaches is to encourage developers to synchronize threads via easy to use \emph{message passing}, namely channels.
However, a study conducted by Tu, Liu, Song and Zhan~\cite{tu2019go} analyzed popular real-world Go projects and their concurrency bugs.
They came to the surprising conclusion that \emph{message passing} produced even more errors than \emph{shared memory}.

A promising solution to atomicity and order violation bugs is \emph{software transactional memory} (STM) as proposed by Peyton Jones~\cite{peytonjones2007beautiful}.
It is an alternative to traditional lock-based synchronization where atomic regions get declared explicitly.
This way STM is one possibility to ensure that there are no side-effects of atomic operations and no data races.
Languages like Haskell provide STM by their design of language but Go can also utilize this mechanism by using external libraries.
Lu, Park, Seo and Zhou~\cite{lu2008mistakes} conducted a study in 2008 where they analyzed the characteristics of real-world concurrency bugs.
One of their key finding was that most of the examined non-deadlock concurrency bugs were covered by atomicity-violation or order-violation patterns.
Their conclusion is that using STM could have avoided about one third of the examined concurrency bugs.

Combining all this with the ongoing trend of using micro-service architectures over big monoliths, these tools could get a lot more useful because the code bases of each service tend to get smaller.
This means that the overhead of instrumenting, testing or model-checking one service gets smaller in absolute numbers.
This way teams can decide more easily which applications need to have a higher concurrency bug coverage and which applications can be ignored.
Record and replay could also be deployed to just some replicas of an instance, reducing the impact on speed whilst collecting useful information to optimize the whole code base.

@@

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}

>>>>

% !TeX encoding = UTF-8
% !TeX program = pdflatex
% !BIB program = biber

\documentclass[english]{lni}

\usepackage[
    backend=biber,
    style=LNI,
]{biblatex}
\addbibresource{references.bib}

\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{float}

@@

\begin{document}
\title{Finding Concurrency Bugs in Production}
\author[Marvin Strangfeld]
{Marvin Strangfeld\footnote{RWTH Aachen University, \email{marvin.strangfeld@rwth-aachen.de}}}
\editor{Gesellschaft für Informatik} % Names of Editors
\booktitle{SKILL 2022} % Name of book title
\yearofpublication{2022}
%%%\lnidoi{18.18420/provided-by-editor-02} % if known
\maketitle

\begin{abstract}
Bugs that are caused by the parallel execution of a program, so called
concurrency bugs are often hard to find and therefore hard to fix.
In this paper we evaluate different techniques and tools that are currently available to effectively find, fix and avoid concurrency bugs.
These tools are evaluated by their ease of use, their computational and storage overhead, their concurrency bug coverage and their potential amount of false-positive reports.
The goal is to find tools that are production-ready and applicable to real-world software projects.
\end{abstract}
\begin{keywords}
    Software testing and Debugging \and
    Concurrent programming languages
\end{keywords}
%%% Beginn des Artikeltexts
\section{Introduction}

@@

\subsection{The Go Programming Language}
The increasing demand for parallel programs has also lead to the development of new programming languages which specialize in this task.
The goal of these languages is to ease the usage of parallel programming patterns, so developers can use them more often without writing extensively more code.
One example for such a language is Go which is a statically typed programming language developed by Google, that compiles down to native binaries~\cite{goDocs}.
Go recently gained a lot of popularity for the purpose of writing highly concurrent programs.
Popular open-source projects like \emph{Kubernetes}\footnote{\url{https://kubernetes.io/}}, \emph{Docker}\footnote{\url{https://www.docker.com/}} and \emph{Prometheus}\footnote{\url{https://prometheus.io/}} are all written in Go.
The official documentation of Go states: ``[... Go's] concurrency mechanisms make it easy to write programs that get the most out of multicore and networked machines [...]''~\cite{goDocs}.
Go abstracts the creation of threads by using lightweight \emph{goroutines}, which can be created by using the \lstinline{go} keyword followed by a named or anonymous function.
The Go runtime maps these goroutines to operating system threads during execution.
Go encourages programmers to use concurrency by \emph{message passing}, which is assumed to be safer and more convenient to use than \emph{shared memory}.
For this purpose Go is using a concept called \emph{channels} where typed messages can be passed through like a UNIX pipeline.
Beside the message passing however, it is also possible to make use of \emph{shared memory}, which gives the developer a lot of freedom to orchestrate the parallel execution of threads.
Even though Go is designed for programming concurrent programs, this does not mean that Go programs are free of concurrency bugs.
Projects written in Go also tend to have more concurrency than projects written in traditional programming languages such as C~\cite{tu2019go}.
So it is even more important for Go developers to catch concurrency bugs effectively and as early as possible.
Together with the increasing popularity of the language, the demand for production-ready tools to catch concurrency bugs is very high, which is why this paper evaluates existing techniques in the context of the Go programming language.

@@

The further structure of this paper continues as follows:
\Cref{sct:taxonomy} gives a brief introduction on the different types of concurrency bugs and their main causes.
\Cref{sct:dynamic} covers some techniques to dynamically detect concurrency bugs and reliably reproduce them.
\Cref{sct:testing} shows different methods of concurrency-aware testing to detect concurrency bugs automatically in a testing environment, where in contrast to a production environment concurrency bugs are explicitly provoked.
Finally, \Cref{sct:static} covers how to detect concurrency bugs with static code analysis.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\linewidth]{figures/ConcurrencyBugClasses.pdf}
    \caption{A Taxonomy of Concurrency Bugs -- based on\cite{tchamgoue2012testing}}
    \label{fig:classes}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[height=0.4\textheight]{figures/Concurrency-Testing.pdf}
    \caption{Automatic concurrency-aware testing for Go programs -- based on\cite{acm2002}}
    \label{fig:testing}
\end{figure}

@@

@@

% \bibliography{references}
\printbibliography
\end{document}
